Tropical Andes IUCN Montane and Lowland forest

### 2023-07-31 code/L0/L0_TropicalAndes_IUCNhabitatClassification.Rmd
update input files to just moist montane and lowland forest crop each to Tropical Andes combine the two resulting rasters save rasters as output files

# Load in Ecosystem Functional Groups
Data downloaded from https://zenodo.org/record/4058819
Citation: Martin Jung, Prabhat Raj Dahal, Stuart H. M. Butchart, Paul F. Donald, Xavier De Lamo, Myroslava Lesiv, Valerie Kapos, Carlo Rondinini, & Piero Visconti. (2020). A global map of terrestrial habitat types (Version 004) [Data set]. Zenodo. https://doi.org/10.5281/zenodo.4058819

```{r}
iucn_habitats <- raster(file.path(data_path, "iucn_habitatclassification_composite_lvl2_ver004/iucn_habitatclassification_composite_lvl2_ver004.tif"))
iucn_habitats_colors <- read.csv(file.path(data_path, "iucn_habitatclassification_composite_lvl2_ver004/color_classification.csv"))
```

```{r}
plot(iucn_habitats)
```

Crop extent - trim to Andean States
```{r}
# Pull in world map
worldMap <- ne_countries(scale = "medium", type = "countries", returnclass = "sf")

@@

```{r}
iucn_habitats_crop <- crop(iucn_habitats, study_region)
TA_iucn_habitats <- mask(iucn_habitats_crop, study_region)
plot(TA_iucn_habitats)
```

value 109 is Forest-Subtropical moist montane & value 106 is Forest - Subtropical moist lowland
see level2.clr or lvl2_style.qml file
```{r}
#rename for masking
TA_iucn_habitats_mask <- TA_iucn_habitats
TA_iucn_habitats_mask[TA_iucn_habitats$iucn_habitatclassification_composite_lvl2_ver004 != 106 | TA_iucn_habitats$iucn_habitatclassification_composite_lvl2_ver004 != 109] <- NA
TA_iucn_habitats_forest <- mask(TA_iucn_habitats, TA_iucn_habitats_mask)

plot(TA_iucn_habitats_forest)
```

# Save raster
```{r}
TA_icunHabitat_path <- file.path(output_path, "TropicalAndes_IUCNHabitat_Forest.GTiff")
writeRaster(TA_iucnHabitats, filename = TA_icunHabitat_path, format = "GTiff", overwrite=TRUE)
```

>>>>

# Load in Ecosystem Functional Groups
Data downloaded from https://zenodo.org/record/4058819
Citation: Martin Jung, Prabhat Raj Dahal, Stuart H. M. Butchart, Paul F. Donald, Xavier De Lamo, Myroslava Lesiv, … Piero Visconti. (2020). A global map of terrestrial habitat types (Version 004) [Data set]. Zenodo. https://doi.org/10.5281/zenodo.4058819

```{r}
iucn_moist_montane_forest <- raster(file.path(data_path, "iucn_habitatclassification_composite_lvl2_ver004/lvl2_frac_1km_ver004/iucn_habitatclassification_fraction_lvl2__109_Forest – Subtropical-tropical moist montane__ver004.tif"))
iucn_moist_lowland_forest <- raster(file.path(data_path, "iucn_habitatclassification_composite_lvl2_ver004/lvl2_frac_1km_ver004/iucn_habitatclassification_fraction_lvl2__106_Forest – Subtropical-tropical moist lowland__ver004.tif"))
```

```{r}
plot(iucn_moist_montane_forest)
plot(iucn_moist_lowland_forest)
```

# Crop extent - trim to Andean States
```{r}
# Pull in world map
worldMap <- ne_countries(scale = "medium", type = "countries", returnclass = "sf")

@@

```{r}
# crop moist montane to tropical andes
iucn_moist_montane_forest_crop <- crop(iucn_moist_montane_forest, study_region)
TA_iucn_moist_montane_forest <- mask(iucn_moist_montane_forest_crop, study_region)
plot(TA_iucn_moist_montane_forest)
TA_iucn_moist_montane_forest
```

```{r}
#crop moist lowland to tropical andes
iucn_moist_lowland_forest_crop <- crop(iucn_moist_lowland_forest, study_region)
TA_iucn_moist_lowland_forest <- mask(iucn_moist_lowland_forest_crop, study_region)
plot(TA_iucn_moist_lowland_forest)
TA_iucn_moist_lowland_forest
```

```{r}
# combine TA_iucn_moist_lowland_forest & TA_iucn_moist_montane_forest into one raster
TA_iucn_habitat_forest <- merge(TA_iucn_moist_lowland_forest, TA_iucn_moist_montane_forest)
plot(TA_iucn_habitat_forest)
TA_iucn_habitat_forest
```

# Save raster
```{r}
writeRaster(TA_iucn_habitat_forest, filename = file.path(output_path, "TropicalAndes_IUCNHabitat_Forest.GTiff"), format = "GTiff", overwrite=TRUE)
writeRaster(TA_iucn_moist_montane_forest, filename = file.path(output_path, "TropicalAndes_IUCNHabitat_Forest_Montane.GTiff"), format = "GTiff", overwrite=TRUE)
writeRaster(TA_iucn_moist_lowland_forest, filename = file.path(output_path, "TropicalAndes_IUCNHabitat_Forest_Lowland.GTiff"), format = "GTiff", overwrite=TRUE)
```
